---
output:
  pdf_document: default
  html_document: default
---
#install.packages("png")

---
title: "A Friendly Intro to R - BioData Club"
author: "Danielle Robinson & Lilly Winfree"
date: "May 26 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Before we begin
- Download R - this is the binary containing R language stuff - it comes with a basic interface, but we won't use it. You need to download this first so your computer will know what to do with R Studio.
    + https://cran.rstudio.com/
- Download R Studio - this is an interface that makes it easy to work with R.
    + https://www.rstudio.com/products/rstudio/download3/
- Go to our Github page: 
    + https://github.com/BioData-Club/Friendly-Intro-to-R
```{r fig.width=5, fig.height=3,echo=FALSE}
library(png)
library(grid)
img <- readPNG("~/Friendly-Intro-to-R/Github.png")
 grid.raster(img)
```
- Click on "Clone or Download"
- Click on "Download Zip"

## Why should you use R??
- Free!
    + unlike Prism, excel
- Open Source 
    + Tons of support, packages, tools
    + Don't trap your file in proprietary format!
- Great for stats
    + Defaults not hidden, question assumptions

## Basics (1)
- use # to comment
- "**<-**" assigns values to a variable
```{r}
x<-1
x
```
- functions ()
```{r}
print(x)
```
## Basics (2)
- help() and ?
- in R console, use up arrow to re-type what you just wrote
- variable names can't start with a number
    + eg Data1 not 1Data
- throwing errors is normal! You will create errors _a lot_

### R is _stupid_ and only knows what _you_ input


##This is what R Studio looks like:
```{r fig.width=8, fig.height=5,echo=FALSE}
library(png)
library(grid)
img <- readPNG("~/Friendly-Intro-to-R/rstudio.png")
 grid.raster(img)
```

## Resources
- CRAN: https://cran.r-project.org
- http://stackoverflow.com - like google for coding questions
- R Markdown cheat sheets (rmarkdown.rstudio.com)
- Ted Laderas' R bootcamp: https://github.com/laderast/r-bootcamp
- Mozilla study groups: https://github.com/mozillascience/studyGroupLessons/issues
- R Open Sci (https://ropensci.org)
- R meetup groups (https://www.meetup.com/portland-r-user-group/)
- additional plotting abilities with ggplot package (we're not going into this)

## Fun Events:
```{r fig.width=8, fig.height=5,echo=FALSE}
library(png)
library(grid)
img <- readPNG("~/Friendly-Intro-to-R/SHD_2017.png")
 grid.raster(img)
```

## Let's go!
From the Github downloaded files, open:

- "example_lengths_for_R.csv"
- "BasicStatsCode.pdf"

We'll start with "BasicStatsCode.pdf" 


###_PLEASE ASK US IF YOU ARE CONFUSED! :-)_


Use your **yellow** sticky note for "confused" and **green** for "ok"


